---
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/jzeng/Desktop/ur/Fall 2022/CSC 262")
```
***5. a)***\
``` {r frame_gender 1}
diabetes <- read.csv("diabetes_fall2022.csv")
table1 <- table(diabetes$gender, diabetes$frame)
addmargins(table1)
```
***5. b)***\
```{r frame_gender 2}
barplot(prop.table(table1,margin=1), xlab="Frame", legend.text = rownames(table1), 
        col=c("darkolivegreen3", "goldenrod3")) 
```
***5. c)***\
$H_0:$ frame and gender are independent vs $H_1:$ frame and gender are associated
$\alpha = 0.05$
```{r frame_gender 3}
chisq.test(table1, correct=F)
qchisq(1-0.05, df = (3-1)*(2-1))
```
We get a critical value of 5.9915 and a p-value of 0.00008548. Since $5.9915 < 18.734 = X^2$ and $0.00008548 < 0.05 = \alpha$, we reject the null hypothesis that frame and gender are independent and conclude that there is an association between frame and gender.\ 
***6. a)***\
$H_0:\sigma_m^2 = \sigma_f^2$ vs $H_1:\sigma_m^2 \ne \sigma_f^2$
$\alpha = 0.01$
```{r stab.glu_gender 1}
s_m <- var(diabetes[diabetes$gender=="male",]$stab.glu)
n_m <- length(diabetes[diabetes$gender=="male",]$stab.glu)
s_f <- var(diabetes[diabetes$gender=="female",]$stab.glu)
n_f <- length(diabetes[diabetes$gender=="female",]$stab.glu)

f_obs <- s_m/s_f
f_obs
2*(1-pf(f_obs, n_m-1, n_f-1))
```
Since p-value = $4.1233*10^{-7} < 0.01 = \alpha$, we reject the null hypothesis that the variances of stabilized glucose for gender are the same.\

***6. b)***\
$H_0:\mu_m - \mu_f \le 0$ vs $H_1:\mu_m - \mu_f > 0$\
$\alpha = 0.05$
```{r stab.glu_gender 2}
t.test(diabetes[diabetes$gender=="male",]$stab.glu, diabetes[diabetes$gender=="female",]$stab.glu, 
       alternative="greater", conf.level = 0.95)
```
Since p-value = $0.0653 > 0.05 = \alpha$, we fail to reject the null hypothesis that the mean stabilized glucose of males is less than or equal to the stabilized glucose of females.\ 
***7. a)***\ 
```{r stab.glu 1}
plot(diabetes$glyhb, diabetes$stab.glu, pch=4, main="Stabilized glucose over glycosylated hemoglobin", 
     ylab = "Stabilized Glucose", xlab = "Glycosylated Hemoglobin")
cor(diabetes$glyhb, diabetes$stab.glu)
```
There appears to be a moderately strong linear relationship between the two variables based on the correlation of  0.7408 and the scatterplor, indicating linear regression to be appropriate.

***7. b)***\ 
$H_0: \rho = 0$ vs $H_1: \rho \ne 0$\ 
$\alpha = 0.05$\ 
```{r stab.glu 2}
cor.test(diabetes$glyhb, diabetes$stab.glu)
qt(1-(0.05/2), df = length(diabetes$glyhb)-2)
```
We find the p-value to be less than $2.2*10^{-16}$ and the test statistic to be 21.186. Since $2.2*10^{-16} < 0.05 = \alpha$, we reject the null hypothesis that stabilized glucose and glycosylated hemoglobin are independent.
***7. c)***\ 
```{r stab.glu 3}
lm1 <- lm(diabetes$stab.glu~diabetes$glyhb)
lm1
```
The regression equation is $y = 17.930x + 6.945$\ 

***7. d)***\ 
$\alpha = 0.01$
```{r stab.glu 4}
confint(lm1, level=0.99)
```
We are 99% confident that the slope for the regression equation falls between 15.7391 and 20.1218.

***7. e)***\
```{r stab.glu 5}
summary(lm1)
```
54.88% of the variability in the glycosylated hemoglobin is explained by its linear relationship with stabilized glucose.

***7. f)***\ 
```{r stab.glu 6}
lm2 <- lm(diabetes$stab.glu~diabetes$glyhb+diabetes$bmi+diabetes$frame)
lm2
```
$y_{large} = 17.7060x_{glyhb} * 0.1715x_{bmi} + 9.3176$\ 
$y_{medium} = 17.7060x_{glyhb} * 0.1715x_{bmi} + 0.0298$\ 
$y_{small} = 17.7060x_{glyhb} * 0.1715x_{bmi} + 3.0768$\ 
The slope for all the equations is the same but each different frame has a different intercept.

***7. g)***\ 
```{r stab.glu 7}
summary(lm2)
```
Multiple R-squared is 0.5548 and the adjusted R-squared is 0.55, since the adjusted R-squred is lower, that means that we added a variable that is not impactful on the output.\ 

***7. h)***\ 
$H_0:$ The bmi variable does not predict stabilized glucose vs $H_1:$ The bmi variable significantly predicts stabilized glucose\ 
$H_0:$ The frame variable does not predict stabilized glucose vs $H_1:$ The frame variable significantly predicts stabilized glucose\ 
$\alpha = 0.05$\ 
```{r stab.glu 8}
lm3 <- lm(diabetes$stab.glu~diabetes$bmi+diabetes$frame)
anova(lm3)
```
We find the test statistic for bmi to be 6.8555 and the test statistic for frame to be 3.6580. Since the bmi p-value = $0.009203 < 0.05 = \alpha$, we reject the null hypothesis that the bmi variable does not predict stabilized glucose. Since the frame p-value  = $0.026726 < 0.05 = \alpha$, we reject the null hypothesis that the frame variable does not predict stabilized glucose.\ 

***7. i)***\ 
```{r stab.glu 9}
par(mfrow=c(2,2))
plot(lm2)
```
In the residuals vs fitted graph, we see increasing variability which indicates that the relationship is potentially non-linear and our regression line is not appropriate for modeling the data. For our homoscedasticity, we do not see a constant variance across all values of x. The QQ plot shows that the data skews right which means that our values do not have a normal distrbution.